Concurrent Programming in Harmony

### 2021-02-01 book/paper.tex
Latest

Harmony can report the following failure types:
\begin{itemize}
\item \texttt{Safety violation}: this means something went wrong with
at least one of the executions of the program that it tried.  This
can include a failing assertion, divide by zero, using an uninitialized
or non-existent variable, dividing a set by an integer, and so on.
It also includes certain infinite loops.
Harmony will print a trace of the shortest bad execution that it found.
\item \texttt{Non-terminating States}: Harmony found one or more states
from which there does not exist an execution such that all threads
terminate.  Harmony will not only print the non-terminating state with
the shortest trace, but also the list of threads
at that state, along with their program counters.
% \item \texttt{Stopped States}: Similar to non-terminating states,
% these are states in which some threads are left that cannot
% terminate.  \autoref{ch:synch} will explain what it means for
% a thread to be stopped.
\end{itemize}

@@

We call a thread \emph{blocked}
\index{blocked thread}%
if a thread cannot change the state or terminate unless
another thread changes the state first.
A thread trying to
acquire a test-and-set spinlock held by another thread is a good example
of a thread being blocked.
The only way forward is if the other thread releases the spinlock.
A thread that is in an infinite loop is also considered blocked,
although Harmony can often detect infinite loops specifically.

@@

Details:
\begin{itemize}
\item The frame pointer points to the current \emph{stack frame},
which consists of the caller's frame pointer and variables, the argument to
the method, an ``invocation type atom'' (\texttt{normal}, \texttt{interrupt}, or
\texttt{thread}), and the return address (in case of \texttt{normal}).
\item A thread terminates when it reaches the \texttt{Return} instruction
of the top-level method (when the stack frame is of type \texttt{thread})
or when it hits an exception.  Exceptions include divide by zero,
reading a non-existent key in a dictionary, accessing a non-existent
variable, infinite loops of microsteps, as well as when an assertion fails;
\item The execution of a thread in \emph{atomic mode} does not get interleaved
with that of other threads.
\item The execution of a thread in \emph{read-only mode} is not allowed
to update shared variables of spawn threads.
\item The register of a thread always contains a dictionary, mapping
atoms to arbitrary values.  The atoms correspond to the variable names
in a Harmony program.
\end{itemize}

>>>>

Harmony can report the following failure types:
\begin{itemize}
\item \texttt{Safety violation}: this means something went wrong with
at least one of the executions of the program that it tried.  This
can include a failing assertion, divide by zero, using an uninitialized
or non-existent variable, dividing a set by an integer, and so on.
% It also includes certain infinite loops.
Harmony will print a trace of the shortest bad execution that it found.
\item \texttt{Non-terminating States}: Harmony found one or more states
from which there does not exist an execution such that all threads
terminate.  Harmony will not only print the non-terminating state with
the shortest trace, but also the list of threads
at that state, along with their program counters.
% \item \texttt{Stopped States}: Similar to non-terminating states,
% these are states in which some threads are left that cannot
% terminate.  \autoref{ch:synch} will explain what it means for
% a thread to be stopped.
\end{itemize}

@@

We call a thread \emph{blocked}
\index{blocked thread}%
if a thread cannot change the state or terminate unless
another thread changes the state first.
A thread trying to
acquire a test-and-set spinlock held by another thread is a good example
of a thread being blocked.
The only way forward is if the other thread releases the spinlock.
A thread that is in an infinite loop is also considered blocked.
% although Harmony can often detect infinite loops specifically.

@@

Details:
\begin{itemize}
\item The frame pointer points to the current \emph{stack frame},
which consists of the caller's frame pointer and variables, the argument to
the method, an ``invocation type atom'' (\texttt{normal}, \texttt{interrupt}, or
\texttt{thread}), and the return address (in case of \texttt{normal}).
\item A thread terminates when it reaches the \texttt{Return} instruction
of the top-level method (when the stack frame is of type \texttt{thread})
or when it hits an exception.  Exceptions include divide by zero,
reading a non-existent key in a dictionary, accessing a non-existent
variable, as well as when an assertion fails;
\item The execution of a thread in \emph{atomic mode} does not get interleaved
with that of other threads.
\item The execution of a thread in \emph{read-only mode} is not allowed
to update shared variables of spawn threads.
\item The register of a thread always contains a dictionary, mapping
atoms to arbitrary values.  The atoms correspond to the variable names
in a Harmony program.
\end{itemize}
